---
title: "Prediction Assignment Writeup"
author: "Timo Olkkonen"
date: "December 21, 2014"
output: html_document
---

Let's load the original data, and remove columns with irrelevant information and/or bad data.

```{r}
originalData <- read.csv("pml-training.csv", header=TRUE)
columnsToBeRemoved <- c("X", "user_name", "raw_timestamp_part_1", "raw_timestamp_part_2", "cvtd_timestamp", "new_window", "num_window", "kurtosis_roll_belt", "kurtosis_picth_belt", "kurtosis_yaw_belt", "kurtosis_yaw_belt", "skewness_roll_belt", "skewness_roll_belt.1", "skewness_yaw_belt", "max_yaw_belt", "min_yaw_belt", "amplitude_yaw_belt", "kurtosis_roll_arm", "kurtosis_picth_arm", "kurtosis_yaw_arm", "skewness_roll_arm", "skewness_pitch_arm", "skewness_yaw_arm", "kurtosis_roll_dumbbell", "kurtosis_picth_dumbbell", "kurtosis_yaw_dumbbell", "skewness_roll_dumbbell", "skewness_pitch_dumbbell", "skewness_yaw_dumbbell", "max_yaw_dumbbell", "min_yaw_dumbbell", "amplitude_yaw_dumbbell", "kurtosis_roll_forearm", "kurtosis_picth_forearm", "kurtosis_yaw_forearm", "skewness_roll_forearm", "skewness_pitch_forearm", "skewness_yaw_forearm", "max_yaw_forearm", "min_yaw_forearm", "amplitude_yaw_forearm", 'max_roll_belt', 'max_picth_belt', 'min_roll_belt', 'min_pitch_belt', 'amplitude_roll_belt', 'amplitude_pitch_belt', 'var_total_accel_belt', 'avg_roll_belt', 'stddev_roll_belt', 'var_roll_belt', 'avg_pitch_belt', 'stddev_pitch_belt', 'var_pitch_belt', 'avg_yaw_belt', 'stddev_yaw_belt', 'var_yaw_belt', 'var_accel_arm', 'avg_roll_arm', 'stddev_roll_arm', 'var_roll_arm', 'avg_pitch_arm', 'stddev_pitch_arm', 'var_pitch_arm', 'avg_yaw_arm', 'stddev_yaw_arm', 'var_yaw_arm', 'max_roll_arm', 'max_picth_arm', 'max_yaw_arm', 'min_roll_arm', 'min_pitch_arm', 'min_yaw_arm', 'amplitude_roll_arm', 'amplitude_pitch_arm', 'amplitude_yaw_arm', 'max_roll_dumbbell', 'max_picth_dumbbell', 'min_roll_dumbbell', 'min_pitch_dumbbell', 'amplitude_roll_dumbbell', 'amplitude_pitch_dumbbell', 'var_accel_dumbbell', 'avg_roll_dumbbell', 'stddev_roll_dumbbell', 'var_roll_dumbbell', 'avg_pitch_dumbbell', 'stddev_pitch_dumbbell', 'var_pitch_dumbbell', 'avg_yaw_dumbbell', 'stddev_yaw_dumbbell', 'var_y', 'var_yaw_dumbbell', 'max_roll_forearm', 'max_picth_forearm', 'min_roll_forearm', 'min_pitch_forearm', 'amplitude_roll_forearm', 'amplitude_pitch_forearm', 'var_accel_forearm', 'avg_roll_forearm', 'stddev_roll_forearm', 'var_roll_forearm', 'avg_pitch_forearm', 'stddev_pitch_forearm', 'var_pitch_forearm', 'avg_yaw_forearm', 'stddev_yaw_forearm', 'var_yaw_forearm')
cleanedData <- originalData[, -which(names(originalData) %in% columnsToBeRemoved)]
```
Now, let's split the data to training and testing sets. Let's use rule-of-thumb-ish 3/4 split.
```{r}
library(caret)
inTrain <- createDataPartition(y=cleanedData$classe, p=0.75, list=FALSE)
training <- cleanedData[inTrain,]
testing <- cleanedData[-inTrain,]
```
Let's try random forest. However, as we have relatively large amount of data, let's restrict number of iterations to three.
```{r, echo=FALSE}
model <- train(classe~., training, method="rf", trControl = trainControl(method = "cv", number = 3, allowParallel = TRUE))
prediction <- predict(model, testing)
```

```{r}
confusionMatrix(testing$classe, prediction)
```
From the output we can see that our model gives pretty good accuracy of 0.9939. We can use this to give predictions in the second part of the homework.
